\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{listings}
%\usepackage{subfig}

\usetheme{Madrid}
\usecolortheme{seahorse}
\usefonttheme{serif} 
 
 
%Information to be included in the title page:
\title{Insertion Sort}
\author{Vivek K. S., Deepak G.}
\institute{Information Systems Decision Sciences (ISDS)\\
MUMA College of Business\\
University of South Florida \\
Tampa, Florida}
\date{2017}
 
 
\begin{document}
\frame{\titlepage}
\begin{frame}
\frametitle{Introduction}
\begin{itemize}
\item The insertion sort has  a different way of working although it still only offers O(n-squared).
\item It maintains a sorted sublist in the lower positions of the list.
\item Each new item is then inserted back into this sublist such that the sorted sublist grows by one item at a time.
\item In the beginning, we start with a list of one item at position 0, which is already sorted.
\item Every pass, through items from position 1 through n-1, compares the current item against those in the sorted sublist.
\item The items that are greater are shifted to the right.
\item When an item that is smaller than the current item is encountered, it can be inserted.
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{How does it work?}
\begin{itemize}
\item Imagine a list [12,24,43,45,64,81] which as we can see is already sorted. This  is the sorted sublist that we have.
\item Lets say we want to insert 44 into this list.
\item The current item 44 will be compared against each item of the sorted sublist.
\item Initially, 81 will be moved to the right as it is larger than 44. 
\item The same happens with 45 and 64.
\item When it encounters 44, which is smaller than 45, it performs an insertion.
\item Animation - http://www.cs.armstrong.edu/liang/animation/web/InsertionSort.html
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Contd}
\begin{itemize}
\item There are n-1 passes to be made for a list of n items again.
\item The iterations starts at 1 and moves through position n-1. Note that position 0 is already sorted when we start.
\item The maximum number of comparisons for an insertion sort is the sum of the first n-1 integers.
\item Again, this is O(n-squared).
\item It is well worth noticing that here, we are shifting instead of exchanging. 
\item In general, shifting requires only a third of the processing power required for exchanging since only one assignment is performed. 
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Python Implementation}
\begin{lstlisting}[language=Python]
def insertion_sort(a_list):
    for index in range(1, len(a_list)):
        current_value = a_list[index]
        position = index
        while position > 0 and 
        a_list[position - 1] > current_value:
            a_list[position] = a_list[position - 1]
            position = position - 1
        a_list[position] = current_value
\end{lstlisting}
Notice that we are not swapping the items but simply assigning them to the target position.
\end{frame}
\end{document}